\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-11 paper.qmd
Set the plot's aspect ratio to 1

<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
```{r}
#| fig-cap: "The $\\gD{}$ layout from tSNE applied for the MNIST digit 1 dataset with defualt (hyper-)parameters. We use our model in the $10\\text{-}D$ space to assess whether this is an accurate representation of data structure present in $10\\text{-}D$, or if it is misleading. The angle of the digit 1 images varies along this structure. Images at the top-right of the $\\gD{}$ layout show the digit 1 angled more to the right, while those at the bottom-right show the digit 1 angled more to the left."
#| label: fig-tsne-best
#| fig-pos: H
#| out-height: 22%

>>>>

<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
```{r}
#| fig-cap: "The $\\gD{}$ layout from tSNE applied for the MNIST digit 1 dataset with defualt (hyper-)parameters. We use our model in the $10\\text{-}D$ space to assess whether this is an accurate representation of data structure present in $10\\text{-}D$, or if it is misleading. The angle of the digit 1 images varies along this structure. Images at the top-right of the $\\gD{}$ layout show the digit 1 angled more to the right, while those at the bottom-right show the digit 1 angled more to the left."
#| label: fig-tsne-best
#| fig-pos: H
#| out-height: 22%
#| eval: false

### 2024-09-11 paper.tex
Set the plot's aspect ratio to 1

The MSE for different binwidths (\(a_1\)) using tSNE, UMAP, PHATE,
PaCMAP, and TriMAP with default (hyper-)parameter setting
(Figure~\ref{fig-mnist-mse}) were calculated. It is found that tSNE
(Figure~\ref{fig-mnist-mse} (a)) provide the most reasonable
representation for the digit 1 dataset, showing universally best.
However, \gD{} representation shows a big non-linear cluster and a small
cluster with a small gap (Figure~\ref{fig-tsne-best}).

In the case of digit 1 data, there should not be any clusters unless
anomalies exist, indicating different digit 1 patterns. The angle of the
digit 1 images varies along this non-linear clustering structure
(Figure~\ref{fig-tsne-best}), while the small cluster contains the digit
1 images with different patterns of the digit 1, unlike the usual
(Figure~\ref{fig-model-error-mnist} (c)). This provides the evidence for
two close clusters.

@@

\includegraphics[width=1\textwidth,height=0.22\textheight]{paper_files/figure-pdf/fig-tsne-best-1.pdf}

\caption{\label{fig-tsne-best}The \(\gD{}\) layout from tSNE applied for
the MNIST digit 1 dataset with defualt (hyper-)parameters. We use our
model in the \(10\text{-}D\) space to assess whether this is an accurate
representation of data structure present in \(10\text{-}D\), or if it is
misleading. The angle of the digit 1 images varies along this structure.
Images at the top-right of the \(\gD{}\) layout show the digit 1 angled
more to the right, while those at the bottom-right show the digit 1
angled more to the left.}

@@

\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the tSNE layout
(a), and three views (b, c, d) of the fit in projections from
\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.07\),
\(b = 456 \  (19,  24)\), \(m = 226\), and benchmark value to remove
large edges is \(0.114\)). (b) The non-linear structure observed in the
\gD{} layout of tSNE (Figure~\ref{fig-tsne-best}) is also visible when
visualizing the model overlaid on the data space. This indicates that
tSNE accurately captures the non-linear structure of the \(10\text{-}D\)
data. (c) the model shows a twisted pattern within the non-linear
structure in \(10\text{-}D\) space, which is an additional pattern not
visible in the \gD{} representation (Figure~\ref{fig-tsne-best}). (d)
some long edges exist in the \(10\text{-}D\) space that are not
recognized as long edges in the \gD{} representation. The full video is
available at \url{https://youtu.be/zcg_GXBmqjA}.}

>>>>

The MSE for different binwidths (\(a_1\)) using tSNE, UMAP, PHATE,
PaCMAP, and TriMAP with default (hyper-)parameter setting
(Figure~\ref{fig-mnist-mse}) were calculated. It is found that tSNE
(Figure~\ref{fig-mnist-mse} (a)) provide the most reasonable
representation for the digit 1 dataset, showing universally best.
However, \gD{} representation shows a big non-linear cluster and a small
cluster with a small gap (\textbf{?@fig-tsne-best}).

In the case of digit 1 data, there should not be any clusters unless
anomalies exist, indicating different digit 1 patterns. The angle of the
digit 1 images varies along this non-linear clustering structure
(\textbf{?@fig-tsne-best}), while the small cluster contains the digit 1
images with different patterns of the digit 1, unlike the usual
(Figure~\ref{fig-model-error-mnist} (c)). This provides the evidence for
two close clusters.

@@

@@

\caption{\label{fig-mnist-tri-proj}Model in \gD{}, on the tSNE layout
(a), and three views (b, c, d) of the fit in projections from
\(10\text{-}D\), for the digit 1 of MNIST data (\(a_1 = 0.07\),
\(b = 456 \  (19,  24)\), \(m = 226\), and benchmark value to remove
large edges is \(0.114\)). (b) The non-linear structure observed in the
\gD{} layout of tSNE (\textbf{?@fig-tsne-best}) is also visible when
visualizing the model overlaid on the data space. This indicates that
tSNE accurately captures the non-linear structure of the \(10\text{-}D\)
data. (c) the model shows a twisted pattern within the non-linear
structure in \(10\text{-}D\) space, which is an additional pattern not
visible in the \gD{} representation (\textbf{?@fig-tsne-best}). (d) some
long edges exist in the \(10\text{-}D\) space that are not recognized as
long edges in the \gD{} representation. The full video is available at
\url{https://youtu.be/zcg_GXBmqjA}.}
